Modelling BCG vaccination in the UK

### 2019-08-08 rmds/05-evidence-policy-change.Rmd
thesis edits + readme

In England and Wales, universal school-aged (13 years old) vaccination was introduced after a MRC trial in the 1950s estimated BCG's effectiveness at 78% in the ethnic White UK born population.[@Hart1972] The policy remained in place until 2005, when England and Wales changed to targeted vaccination of high risk neonates. The 2005 change in BCG vaccination policy was motivated by evidence of decreased transmission of TB, an increasing proportion of TB cases occurring in the non-UK born,[@PHE2017] and modelling evidence that suggested stopping the BCG schools scheme would have minimal long term effects on incidence rates.[@Sutherland1989] Due to the complex nature of both TB and the BCG vaccine, the ongoing impact of this change in policy is hard to directly estimate, with decision makers relying on expert opinion, evidence from surveillance data, and insight from modelling studies.

@@

Using the above information and the projected differences between vaccinated and unvaccinated notificationrates allows the number of prevented notifications for each age group to be found for each year. This can then be combined to give the total number of prevented notifications for those aged between 15-29. To understand these estimates, estimates of the projected yearly notifications if the scheme continues are required. These totals were derived from the vaccinated and unvaccinated rates supplemented with similar projections from the tuberculin positive or otherwise ineligible sourced from the 1983 BCG survey.[@Sutherland1987a]

@@

For each scenario, I ran the model until 2028 with 10,000 parameter samples. I tested the difference between scenarios using the Mann-Whitney test for both the number of vaccines needed to prevent a single case in 15 years after vaccination for a cohort aged 13-14 years old at vaccination and the total number of additional notifications caused by ending the BCG schools scheme. As in [@Sutherland1989] a 15 year time horizon was used with 5 year intervals. The year closest to the year of the change in vaccination policy (2005), which had model estimates, was used as the baseline. The code for this analysis is available online^[Code: https://github.com/seabbs/AssessBCGPolicyChange]

>>>>

In England and Wales, universal school-aged (at 13-14 years old) vaccination was introduced after a MRC trial in the 1950s estimated BCG's effectiveness at 78% in the ethnic White UK born population.[@Hart1972] The policy remained in place until 2005, when England and Wales changed to targeted vaccination of high risk neonates. The 2005 change in BCG vaccination policy was motivated by evidence of decreased transmission of TB, an increasing proportion of TB cases occurring in the non-UK born,[@PHE2017] and modelling evidence that suggested stopping the BCG schools scheme would have minimal long term effects on incidence rates.[@Sutherland1989] Due to the complex nature of both TB and the BCG vaccine, the ongoing impact of this change in policy is hard to directly estimate, with decision makers relying on expert opinion, evidence from surveillance data, and insight from modelling studies.

@@

Using this information and the projected differences between vaccinated and unvaccinated notification rates allows the number of prevented notifications for each age group to be found for each year. This can then be combined to give the total number of prevented notifications for those aged between 15-29. To understand these estimates, estimates of the projected yearly notifications if the scheme continues are required. These totals were derived from the vaccinated and unvaccinated rates supplemented with similar projections from the tuberculin positive or otherwise ineligible sourced from the 1983 BCG survey.[@Sutherland1987a]

@@

For each scenario, I ran the model until 2028 with 10,000 parameter samples. I tested the difference between scenarios using the Mann-Whitney test for both the number of vaccines needed to prevent a single case in 15 years after vaccination for a cohort aged 13-14 years old at vaccination and the total number of additional notifications caused by ending the BCG schools scheme. As in [@Sutherland1989] a 15 year time horizon was used with 5 year intervals. The year closest to the year of the change in vaccination policy (2005), which had model estimates, was used as the baseline. The code for this analysis is available online.^[Code: https://github.com/seabbs/AssessBCGPolicyChange]

### 2019-08-08 rmds/10-model-results.Rmd
wordcount
thesis edits + readme

## Wordcount: 2293

@@

```{r 10-table-future-incidence}
sum_scenario %>% filter(Year %% 5 == 0) %>% 
  set_names(., nm = names(.) %>% str_replace_all("BCG", "BCG (95% CI)")) %>% 
  select(-Observation) %>% 
kable(
  caption.short = "Forecast overall TB incidence for each scenario evaluated from 2005 to 2040.", 
  caption = "Forecast overall TB incidence for each scenario evaluated from 2005 to 2040. For brevity only 5 year intervals are shown. (95\\% CI): 95\\% credible interval estimated using the 2.5\\% quantile and the 97.5\\% quantile.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(latex_options = c("hold_position"))
```

@@

```{r 10-age-scenarios-fig, fig.scap = "Forecast TB incidence for each scenario evaluated from 2005 to 2040, stratified by age group.", fig.cap = "Forecast TB incidence for each scenario evaluated from 2005 to 2040, stratified by age group (0-11). 0-9 refers to 5 year age groups from 0-4 years old to 45-49 years old. 10 refers to those aged between 50 and 69 and 11 refers to those aged 70+. Scenarios are differentiated by colour. The darker ribbon for each colour identifies the interquartile range, whilst the lighter ribbon indicates the 2.5\\% and 97.5\\% quantiles. The line represents the median."}
knitr::include_graphics(file.path(plots_path, "overview-scenarios-future-age-1.png"))
```

@@

```{r 10-table-deaths-preds}
sum_deaths_scenario %>% filter(Year %% 5 == 0) %>% 
  set_names(., nm = names(.) %>% str_replace_all("BCG", "BCG (95% CI)")) %>% 
  select(-Observation) %>% 
kable(
  caption.short = "Forecast overall TB mortality for each scenario evaluated from 2005 to 2040.", 
  caption = "Forecast overall TB mortality for each scenario evaluated from 2005 to 2040. For brevity only 5 year intervals are shown.(95\\% CI): 95\\% credible interval estimated using the 2.5\\% quantile and the 97.5\\% quantile.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(latex_options = c("hold_position"))
```

```{r 10-age-scenarios-deaths-fig, fig.scap = "Forecast TB mortality for each scenario evaluated from 2005 to 2040, stratified by age group, stratified by age group.", fig.cap = "Forecast TB mortality for each scenario evaluated from 2005 to 2040, stratified by age group (0-11). 0-9 refers to 5 year age groups from 0-4 years old to 45-49 years old. 10 refers to those aged between 50 and 69 and 11 refers to those aged 70+. Scenarios are differentiated by colour. The darker ribbon for each colour identifies the interquartile range, whilst the lighter ribbon indicates the 2.5\\% and 97.5\\% quantiles. The line represents the median."}
knitr::include_graphics(file.path(plots_path, "overview-scenarios-deaths-future-age-1.png"))
```

>>>>

## Wordcount: 2292

@@

```{r 10-table-future-incidence}
sum_scenario %>% filter(Year %% 5 == 0) %>% 
  set_names(., nm = names(.) %>% str_replace_all("BCG", "BCG (95% CI)")) %>% 
  select(-Observation) %>% 
kable(
  caption.short = "Forecast of overall TB incidence for each scenario evaluated from 2005 to 2040.", 
  caption = "Forecast of overall TB incidence for each scenario evaluated from 2005 to 2040. For brevity only 5 year intervals are shown. (95\\% CI): 95\\% credible interval estimated using the 2.5\\% quantile and the 97.5\\% quantile.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(latex_options = c("hold_position"))
```

@@

```{r 10-age-scenarios-fig, fig.scap = "Forecast of TB incidence for each scenario evaluated from 2005 to 2040, stratified by age group.", fig.cap = "Forecast of TB incidence for each scenario evaluated from 2005 to 2040, stratified by age group (0-11). 0-9 refers to 5 year age groups from 0-4 years old to 45-49 years old. 10 refers to those aged between 50 and 69 and 11 refers to those aged 70+. Scenarios are differentiated by colour. The darker ribbon for each colour identifies the interquartile range, whilst the lighter ribbon indicates the 2.5\\% and 97.5\\% quantiles. The line represents the median."}
knitr::include_graphics(file.path(plots_path, "overview-scenarios-future-age-1.png"))
```

@@

```{r 10-table-deaths-preds}
sum_deaths_scenario %>% filter(Year %% 5 == 0) %>% 
  set_names(., nm = names(.) %>% str_replace_all("BCG", "BCG (95% CI)")) %>% 
  select(-Observation) %>% 
kable(
  caption.short = "Forecast of overall TB mortality for each scenario evaluated from 2005 to 2040.", 
  caption = "Forecast of overall TB mortality for each scenario evaluated from 2005 to 2040. For brevity only 5 year intervals are shown.(95\\% CI): 95\\% credible interval estimated using the 2.5\\% quantile and the 97.5\\% quantile.", booktabs = TRUE, longtable = TRUE) %>% 
  kable_styling(latex_options = c("hold_position"))
```

```{r 10-age-scenarios-deaths-fig, fig.scap = "Forecast of TB mortality for each scenario evaluated from 2005 to 2040, stratified by age group, stratified by age group.", fig.cap = "Forecast of TB mortality for each scenario evaluated from 2005 to 2040, stratified by age group (0-11). 0-9 refers to 5 year age groups from 0-4 years old to 45-49 years old. 10 refers to those aged between 50 and 69 and 11 refers to those aged 70+. Scenarios are differentiated by colour. The darker ribbon for each colour identifies the interquartile range, whilst the lighter ribbon indicates the 2.5\\% and 97.5\\% quantiles. The line represents the median."}
knitr::include_graphics(file.path(plots_path, "overview-scenarios-deaths-future-age-1.png"))
```
